% !TeX spellcheck = en_GB
\newcommand{\code}{\texttt}
\chapter{Introduction}
\label{sec:introduction}

\epigraph{Every accomplishment starts with the decision to try.}{unknown}
In this chapter, we present the motivation and legitimisation of our thesis and highlight the scope delimitations.

\section{Motivation}
In this first section, we shall legitimate this thesis and explain the value and applicability of our proposed solution.

\subsection{Present Situation}
The Internet Engineering Task Force (\gls{ietf}) standard draft \emph{Using \gls{xmpp} for Security Information Exchange} \cite{ietf-mile-xmpp-grid-05}, as summarised in section~\fullref{sec:ietf-standard-draft-using-xmpp-for-security-information-exchange}, defines a protocol to exchange security-relevant information between network endpoints.
The draft was created by the Managed Incident Lightweight Exchange (\gls{mile}) working group to support computer and network security incident management.

Hereafter, we refer to this \gls{ietf} standard draft as \gls{xmpp-grid-standard}.

To demonstrate the viability of the \gls{xmpp-grid-standard} a rapid prototype was developed in November~2017~\cite{xmpp-grid-prototype}.

\subsection{Problem and Vision}
Currently, no implementation of the \gls{xmpp-grid-standard} management functionality exists that is ready for production use regarding usability and security.

To solve this problem, a graphical interface with bindings to a suitable \gls{broker} must be proposed and implemented.
The interface should permit network administrators to manage and review \glspl{topic}, \glspl{persisted-item} and \glspl{platform}.
Additionally, \glspl{consumer} and \glspl{provider} permissions of \glspl{topic} and \glspl{platform} must be manageable.

The graphical interface supports administrators to better understand the underlying hierarchy and affiliations of \glspl{topic}, enabling them to assess security implications.
As the interface uses familiar terminology known to an administrator, no in-depth understanding of the underlying \gls{xmpp} technology is required to configure and comprehend an \gls{xmpp-grid}.
Finally, the interface also provides better usability than existing command line interfaces and \gls{xmpp} configuration software, which leads to fewer configuration mistakes and improved efficiency.

We hope that with the help of our implementation the \gls{ietf} draft ``Using \gls{xmpp} for Security Information Exchange'' will become an established security standard used in practical industry applications.

\section{Scope Delimitation} % Abgrenzung
As described in the \nameref{sec:task-description}, the focus of this thesis is on the evaluation, design and implementation of the \gls{xmpp-grid-broker}.
Adding missing functionality or fixing complex bugs in existing server or client implementations are beyond the scope of this thesis.
